---
title: "New recipes analysis"
Author: "Tommaso"
output: md_document
---

# Import libraries
```{r include=FALSE}
library(naniar)
library(tidyr)
library(dplyr)
```

# Collect data

### Vegetarian
```{r}
dta_vege <- read.csv("/home/tommaso/diversite-phylogenetique-de-l-assiette-au-jardin/results/Vegetarian/vegetarian_results.tsv", header=TRUE, sep="\t", fill=T)
summary(dta_vege)
```

### Top
```{r}
dta_top <- read.csv("/home/tommaso/diversite-phylogenetique-de-l-assiette-au-jardin/results/Top/top_results.tsv", header=TRUE, sep="\t", fill=T)
summary(dta_top)
```

### Dessert
```{r}
dta_dessert <- read.csv("/home/tommaso/diversite-phylogenetique-de-l-assiette-au-jardin/results/Dessert/dessert_results.tsv", header=TRUE, sep="\t", fill=T)
summary(dta_dessert)
```

# Removing "g" from "dry_matter" column
```{r}
remove_grams <- function (d) {
  d$Dry_matter <- as.numeric(sapply(d$Dry_matter, sub, pattern=" g", replacement=""))
  return(d)
}
dta_vege <- remove_grams(dta_vege)
dta_top <- remove_grams(dta_top)
dta_dessert <- remove_grams(dta_dessert)
```

# Trim and grouping by recipe
```{r}
trim_group_data <- function (d) {
  return(
    subset(d, Richness > 2) %>% group_by(Recipe)
  )
}

vege_by_recipe <- trim_group_data(dta_vege)
top_by_recipe <- trim_group_data(dta_top)
dessert_by_recipe <- trim_group_data(dta_dessert)
```

# Summarize of recipes
```{r}
summarise_by_mean <- function (d) {
  return( d %>% summarise(
    Water = mean(x=Water, na.rm = T), 
    Glucides = weighted.mean(x=Glucides, w=Dry_matter, na.rm = T),
    Lipids = weighted.mean(x=Lipids, w=Dry_matter, na.rm = T),
    Proteins = weighted.mean(x=Proteins, w=Dry_matter, na.rm = T),
    Energy =  weighted.mean(x=Energy, w=Dry_matter, na.rm = T), 
    Phylogenetic_diversity = mean(Phylogenetic_diversity, na.rm = T),
    Weighted_phylogenetic_diversity = mean(Weighted_phylogenetic_diversity, na.rm = T),
    Richness = mean(Richness, na.rm = T),
    Shannon = mean(Shannon, na.rm = T),
    Simpson = mean(Simpson, na.rn = T)
  )
  )
}

vege_simple <- summarise_by_mean(vege_by_recipe)
vege_simple$type <- "Vegetarian"
top_simple <- summarise_by_mean(top_by_recipe)
top_simple$type <- "Top"
dessert_simple <- summarise_by_mean(dessert_by_recipe)
dessert_simple$type <- "Dessert"
```

# Histograms

### Phylogenetic diversity
```{r}
par(mfrow=c(3,1), mar=c(4,3,0,0))
bins <- seq(0, 500, by=50)
hist(vege_simple$Phylogenetic_diversity, ylim=c(0, 15), xlim=c(50, 350), main="", xlab="Vegetarian", col=rgb(0,1,0,0.3), breaks=bins, axes=F)
axis(side=1, at=bins, labels=bins)
axis(side=2, at=c(0,5,10,15), labels=c(0,5,10,15))
hist(top_simple$Phylogenetic_diversity, col=rgb(1, 0, 0, 0.3), ylim=c(0, 15), xlim=c(50, 350), main="", xlab="Top", breaks=bins, axes=F)
axis(side=1, at=bins, labels=bins)
axis(side=2, at=c(0,5,10,15), labels=c(0,5,10,15))
hist(dessert_simple$Phylogenetic_diversity, col=rgb(0, 0, 1, 0.3), ylim=c(0, 15), xlim=c(50, 350), main="", xlab="Desserts", breaks=bins, axes=F)
axis(side=1, at=bins, labels=bins)
axis(side=2, at=c(0,5,10,15), labels=c(0,5,10,15))

```

### Weighted phylogenetic diversity
```{r}
par(mfrow=c(3,1), mar=c(4,3,0,0))
bins <- seq(0, 50, by=5)
hist(vege_simple$Weighted_phylogenetic_diversity, ylim=c(0, 15), xlim=c(0, 60), main="", xlab="Vegetarian", col=rgb(0,1,0,0.3), breaks=bins, axes=F)
axis(side=1, at=bins, labels=bins)
axis(side=2, at=c(0,5,10,15), labels=c(0,5,10,15))
hist(top_simple$Weighted_phylogenetic_diversity, col=rgb(1, 0, 0, 0.3), ylim=c(0, 15), xlim=c(0, 60), main="", xlab="Top", breaks=bins, axes=F)
axis(side=1, at=bins, labels=bins)
axis(side=2, at=c(0,5,10,15), labels=c(0,5,10,15))
hist(dessert_simple$Weighted_phylogenetic_diversity, col=rgb(0, 0, 1, 0.3), ylim=c(0, 15), xlim=c(0, 60), main="", xlab="Desserts", breaks=bins, axes=F)
axis(side=1, at=bins, labels=bins)
axis(side=2, at=c(0,5,10,15), labels=c(0,5,10,15))
```

### Richness
```{r}
par(mfrow=c(3,1), mar=c(4,3,0,0))
bins <- seq(0, 20, by=2)
hist(vege_simple$Richness, ylim=c(0, 20), xlim=c(0, 20), main="", xlab="Vegetarian", col=rgb(0,1,0,0.3), breaks=bins, axes=F)
axis(side=1, at=bins, labels=bins)
axis(side=2, at=c(0,10,20), labels=c(0,10,20))
hist(top_simple$Richness, col=rgb(1, 0, 0, 0.3), ylim=c(0, 20), xlim=c(0, 20), main="", xlab="Top", breaks=bins, axes=F)
axis(side=1, at=bins, labels=bins)
axis(side=2, at=c(0,10,20), labels=c(0,10,20))
hist(dessert_simple$Richness, col=rgb(0, 0, 1, 0.3), ylim=c(0, 20), xlim=c(0, 20), main="", xlab="Desserts", breaks=bins, axes=F)
axis(side=1, at=bins, labels=bins)
axis(side=2, at=c(0,10,20), labels=c(0,10,20))
```

### Shannon
```{r}
par(mfrow=c(3,1), mar=c(4,3,0,0))
bins <- seq(-10, 5, by=1)
hist(vege_simple$Shannon, ylim=c(0, 20), xlim=c(-10, 5), main="", xlab="Vegetarian", col=rgb(0,1,0,0.3), breaks=bins, axes=F)
axis(side=1, at=bins, labels=bins)
axis(side=2, at=c(0,10,20), labels=c(0,10,20))

hist(top_simple$Shannon, col=rgb(1, 0, 0, 0.3), ylim=c(0, 15), xlim=c(-10, 5), main="", xlab="Top", breaks=bins, axes=F)
axis(side=1, at=bins, labels=bins)
axis(side=2, at=c(0,10,20), labels=c(0,10,20))

hist(dessert_simple$Shannon, col=rgb(0, 0, 1, 0.3), ylim=c(0, 15), xlim=c(-10, 5), main="", xlab="Desserts", breaks=bins, axes=F)
axis(side=1, at=bins, labels=bins)
axis(side=2, at=c(0,10,20), labels=c(0,10,20))
```

### Simpson
```{r}
par(mfrow=c(3,1), mar=c(4,3,0,0))
bins <- seq(0, 30, by=2)
hist(vege_simple$Simpson, ylim=c(0, 30), xlim=c(0, 30), main="", xlab="Vegetarian", col=rgb(0,1,0,0.3), breaks=bins, axes=F)
axis(side=1, at=bins, labels=bins)
axis(side=2, at=c(0,10,20,30), labels=c(0,10,20,30))

hist(top_simple$Simpson, col=rgb(1, 0, 0, 0.3), ylim=c(0, 30), xlim=c(0, 30), main="", xlab="Top", breaks=bins, axes=F)
axis(side=1, at=bins, labels=bins)
axis(side=2, at=c(0,10,20,30), labels=c(0,10,20,30))

hist(dessert_simple$Simpson, col=rgb(0, 0, 1, 0.3), ylim=c(0, 30), xlim=c(0, 30), main="", xlab="Desserts", breaks=bins, axes=F)
axis(side=1, at=bins, labels=bins)
axis(side=2, at=c(0,10,20,30), labels=c(0,10,20,30))
```

# Boxplot

### Phylognetic diversity
```{r}
par(mfrow=c(1,1), mar=c(4,3,0,0))
boxplot(vege_simple$Phylogenetic_diversity, top_simple$Phylogenetic_diversity, dessert_simple$Phylogenetic_diversity, col=c("green", "red", "blue"), names=c("Vegetarien", "top", "dessert"))
```

### Weighted phylogenetic diversity
```{r}
par(mfrow=c(1,1), mar=c(4,3,0,0))
boxplot(vege_simple$Weighted_phylogenetic_diversity, top_simple$Weighted_phylogenetic_diversity, dessert_simple$Weighted_phylogenetic_diversity, col=c("green", "red", "blue"), names=c("Vegetarien", "top", "dessert"))
```

### Richness
```{r}
par(mfrow=c(1,1), mar=c(4,3,0,0))
boxplot(vege_simple$Richness, top_simple$Richness, dessert_simple$Richness, col=c("green", "red", "blue"), names=c("Vegetarien", "top", "dessert"))
```

### Shannon
```{r}
par(mfrow=c(1,1), mar=c(4,3,0,0))
boxplot(vege_simple$Shannon, top_simple$Shannon, dessert_simple$Shannon, col=c("green", "red", "blue"), names=c("Vegetarien", "top", "dessert"))
```

### Simpson
```{r}
par(mfrow=c(1,1), mar=c(4,3,0,0))
boxplot(vege_simple$Simpson, top_simple$Simpson, dessert_simple$Simpson, col=c("green", "red", "blue"), names=c("Vegetarien", "top", "dessert"))
```

# Combining recipes
```{r}
all_simple <- rbind(vege_simple, top_simple, dessert_simple)
all_simple$type <- as.factor(all_simple$type)
```

### Plot wheighted phylogenetic diversity by nutritional values
```{r}

par(mfrow=c(3,2), mar=c(4,3,0,0))

plot(all_simple$Energy, all_simple$Weighted_phylogenetic_diversity, pch=20, col=factor(all_simple$type), main="", xlab="Energy")
abline(lm(all_simple$Weighted_phylogenetic_diversity ~ all_simple$Energy), lty=2, col="grey")

plot(all_simple$Water, all_simple$Weighted_phylogenetic_diversity, pch=20, col=factor(all_simple$type), main="", xlab="Water")
abline(lm(all_simple$Weighted_phylogenetic_diversity ~ all_simple$Glucides), lty=2, col="grey")

plot(all_simple$Glucides, all_simple$Weighted_phylogenetic_diversity, pch=20, col=factor(all_simple$type), main="", xlab="Glucides")
abline(lm(all_simple$Weighted_phylogenetic_diversity ~ all_simple$Lipids), lty=2, col="grey")

plot(all_simple$Proteins, all_simple$Weighted_phylogenetic_diversity, pch=20, col=factor(all_simple$type), ylab="Diversité phylogénétique pondérée", main="", xlab="Proteins")
abline(lm(all_simple$Weighted_phylogenetic_diversity ~ all_simple$Proteins), lty=2, col="grey")

plot(all_simple$Lipids, all_simple$Weighted_phylogenetic_diversity, pch=20, col=factor(all_simple$type), main="", xlab="Lipids")
abline(lm(all_simple$Weighted_phylogenetic_diversity ~ all_simple$Lipids), lty=2, col="grey")

plot.new()
legend(x="bottomright", legend=unique(all_simple$type), fill=c("green", "red", "black"))
```

### Correlation with energy
```{r}
cor.test(all_simple$Energy, all_simple$Phylogenetic_diversity)
cor.test(all_simple$Energy, all_simple$Weighted_phylogenetic_diversity)
```

### Correlation with watere
```{r}
cor.test(all_simple$Water, all_simple$Phylogenetic_diversity)
cor.test(all_simple$Water, all_simple$Weighted_phylogenetic_diversity)
```

### Correlation with sugars
```{r}
cor.test(all_simple$Glucides, all_simple$Phylogenetic_diversity)
cor.test(all_simple$Glucides, all_simple$Weighted_phylogenetic_diversity)
```

### Correlation with lipides
```{r}
cor.test(all_simple$Lipids, all_simple$Phylogenetic_diversity)
cor.test(all_simple$Lipids, all_simple$Weighted_phylogenetic_diversity)
```

### Correlation with proteins
```{r}
cor.test(all_simple$Proteins, all_simple$Phylogenetic_diversity)
cor.test(all_simple$Proteins, all_simple$Weighted_phylogenetic_diversity)
```



# Linear model: can we explain phylogenetic diversity based on composition?
```{r}
linear_model <- lm(Phylogenetic_diversity ~ Glucides + Proteins + Lipids, data=all_simple)
summary(linear_model)
```

```{r} 
linear_model.p <- lm(Weighted_phylogenetic_diversity ~ Energy + Water + Glucides + Proteins + Lipids + type, data=all_simple)
summary(linear_model.p)
```
```{r}
linear_model.p <- lm(Phylogenetic_diversity ~ Energy + Water + Glucides + Proteins + Lipids + type, data=all_simple)
summary(linear_model.p)
```


```{r}
keep <- subset(all_simple, select=c("Weighted_phylogenetic_diversity", "Energy", "Water", "Glucides", "Lipids", "Proteins"))
plot(keep)
linear_model <- lm(Weighted_phylogenetic_diversity ~Water + Glucides + Proteins + Lipids + Energy + type, data=all_simple)
summary(linear_model)

plot(linear_model)
```

```{r}
linear_model <- lm(Weighted_phylogenetic_diversity ~ type, data=all_simple)
summary(linear_model)

plot(linear_model)
```


# Link between phylogenetic diversity and richness
```{r}
plot(all_simple$Richness, all_simple$Phylogenetic_diversity, pch=20, col=factor(all_simple$type), ylab="Diversité Phylogénétique", xlab="Richesse"  )
abline(lm(vege_simple$Phylogenetic_diversity ~ vege_simple$Richness), col="green")
#points(top_simple$Richness, top_simple$Phylogenetic_diversity, pch=20, col="red" )
abline(lm(top_simple$Phylogenetic_diversity ~ top_simple$Richness), col="red")
#points(dessert_simple$Richness, dessert_simple$Phylogenetic_diversity, pch=20, col="blue")
abline(lm(dessert_simple$Phylogenetic_diversity ~ dessert_simple$Richness), col="black")
```

# Link between weighted phylogenetic diversity and richness
```{r}
plot(all_simple$Richness, all_simple$Weighted_phylogenetic_diversity, pch=20, col=factor(all_simple$type), ylab="Diversité phylogénétique pondérée", xlab="Richesse")
abline(lm(vege_simple$Weighted_phylogenetic_diversity ~ vege_simple$Richness), col="green")
#points(top_simple$Richness, top_simple$Weighted_phylogenetic_diversity, pch=20, col="red" )
abline(lm(top_simple$Weighted_phylogenetic_diversity ~ top_simple$Richness), col="red")
#points(dessert_simple$Richness, dessert_simple$Weighted_phylogenetic_diversity, pch=20, col="blue")
abline(lm(dessert_simple$Weighted_phylogenetic_diversity ~ dessert_simple$Richness), col="black")
```

# Other plots

### Vegetarian
```{r}
plot(vege_simple$Recipe, vege_simple$Weighted_phylogenetic_diversity, las=2)
```

### Top
```{r}
plot(top_simple$Recipe, top_simple$Weighted_phylogenetic_diversity, las=2)
```

### Dessert
```{r}
plot(dessert_simple$Recipe, dessert_simple$Weighted_phylogenetic_diversity, las=2)
```

## All samples
```{r}
plot(all_simple$Weighted_phylogenetic_diversity, col=factor(all_simple$type), ylab="Weighted phylogenetic diversity")
legend(x="bottomright", legend=unique(all_simple$type), fill=c("green", "red", "black"))
```


